### 2025-04-27 thesis/thesis.tex
Write a chapter about limitations

\documentclass{article}

\usepackage[a4paper]{geometry}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{float}
\usepackage{hyperref}
\usepackage{titlesec}
\usepackage[natbibapa]{apacite}

\graphicspath{{./images/}}

\pagestyle{fancyplain}
\fancyhf{}
\lhead{\fancyplain{}{Mara Schulke} }
\rhead{\fancyplain{}{\today}}
\cfoot{\fancyplain{}{\thepage}}

\newcommand{\figuresource}[1]{
	\begin{center}Quelle: {#1}\end{center}
}

\titleformat{\chapter}{\normalfont\Large\bfseries}{\thechapter.}{20pt}{\Large}
\titleformat{\section}{\normalfont\large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\normalfont\normalsize\bfseries}{\thesubsection}{1em}{}
\titleformat{\subsubsection}{\normalfont\small\bfseries}{\thesubsubsection}{1em}{}

\begin{document}

\begin{titlepage}
    \begin{center}
        \begin{Large}
            Brandenburg University of Applied Sciences \\[1em]
        \end{Large}
        IT Security \\
        Computerscience \\
        Prof. Dr. Oleg Lobachev \\
        MSc. Florian Eich
    \end{center}

    \vfill

    \begin{center}
        \Large{Reliable Natural Language Interfaces using LLMs, Self-Correction and Incremental Schema Analysis}\\[0.5em]
        \large{Bachelor Thesis}\\[1em]

        \begin{normalsize}
            Summer semester 2025\\[0.25em]
            \today
        \end{normalsize}
    \end{center}

    \vfill

    \begin{center}
        Mara Schulke – Matr-Nr. 20215853
    \end{center}
\end{titlepage}

\begin{abstract}
This thesis explores the integration of large language models (LLMs) into PostgreSQL database systems in order
to make the database accessible via natural language instead of the postgres SQL dialect. The research focuses
on implementation strategies, performance optimization, and practical applications of this concept.
\end{abstract}

\tableofcontents

\listoffigures

\section*{List of Abbreviations}
\begin{tabular}{ll}
GPT & Generative Pretrained Transformer \\
SQL & Structured Query Language \\
API & Application Programming Interface \\
LLM & Large Language Model \\
DBMS & Database Management System \\
NL2SQL & Natural Language to SQL \\
\end{tabular}

\newpage

% 1/3 Vorarbeit, Literature Review
% 1/3 Decomp und System Design
% 1/3 Implementation und Evaluation

% Introduction
\section{Introduction}
\subsection{Problem Statement and Motivation}

Database systems represent a backbone of modern computer science, allowing for rapid advancements
whilst shielding us from the problem categories that come along with managing and querying large amounts
of, usually structured, data efficiently. However, most Database Management Systems (DBMS) have
traditionally required specialized knowledge, usually of the Structured Query Language (SQL), in order
to become useable. Whilst this barrier may be percieved differently across diverse usergroups it
represents a fundamental misalignment between end-user goals (e.g. analysts, researchers, domain experts
etc.) and the underlying DBMS, thus often requiring software engineering efforts in order to reduce this friction.

This barrier is the reason entire classes of software projects exists (for example, admin / support panels),
data analytics tools etc. which therefore introduce significant churn and delay between the implementation
of a database system and reaching the desired end user impact. Often these projects span multiple years, require
costly staffing and yield little to no novel technical value.

Emerging technologies such as Large Language Models (LLMs) have proven themselves as a sensible tool for bridging
fuzzy user provided input into discrete, machine readable formats. Prominent models in this field have demostrated
outstanding capabilities that enable computer scientists to tackle new problem classes, that used to be
challenging / yielded unsatisfying results with discrete programming approaches.

This thesis is exploring ways to overcome the above outlined barrier using natural language queries, so that domain experts,
business owners, support staff etc. are able to seamlessly interact with their data, essentially eliminating the
requirement of learning SQL (and its pitfalls). By translating natural language to SQL using Large Language Models
this translation becomes very robust (e.g. against different kinds of phrasing) and enables novel applications
in how businesses, researchers and professionals interact with their data — it represents a fundamental shift 
(ie. moving away from SQL) towards a more inclusive and data driven world.

\subsection{Objectives of the Thesis}

This thesis aims to address the aforementioned challanges when it comes to database accessibility.
The following objectives are the core research area of this thesis:

\begin{enumerate}
    \item Develop a database extension that can translate natural language queries into semantically
          accurate SQL queries using Large Language Models.
    \item To evaluate the effectiveness and feasibility of different Models aswell as prompt engineering
          techniques in order to improve the performance of the system.
    \item Identify and address issues when it comes to handling amibguous, complex and domain specific user input.
    \item Benchmark the performance of the implementation against common natural language to SQL (NL2SQL) benchmarks.
    \item Idenitfy potential use cases for real world scenarios that could deliver a noticable upsides to users.
    \item Analyze the short commings and limitations of this approach and propose potential solutions to overcome them.
\end{enumerate}

\subsection{Research Questions}

% \subsubsection{Primary Research Questions}

\subsubsection*{RQ1 — Are natural language database interfaces feasible for real world application?}

The primary research questions when it comes to natural language database interfaces evolve around their
semantic accuracy and reliability, therefore questioning their feasibility for real world usage.
LLMs have notoriously been known for their ability to hallucinate / produce false, but promising outputs.
This behaviour can be especially  dangerous when opting for data driven decisions that rely on false data
due to a mistranslation from natural language to SQL. LLMs could cause hard to understand and debug behaviour,
like false computation of distributions when the intermediate format is not being shown to the user. This
thesis tries to determine whether such hallucinations could be reasonably prevented and whether the associated
performance and hardware requirements are suitable for a real world deployment, outside of research situations.

Specifically the two big underlying questions are:

\begin{enumerate}
    \item Is the semantic accuracy of natural language database interfaces high enough to yield a noticable
          benefit to users?
    \item Is it possible to run such an interface on reasonable, mass available hardware (e.g. excluding high end research GPUs).
\end{enumerate}

\subsubsection*{RQ2 — What approaches are most effective in resolving ambiguity when translating natural language queries into SQL?}

To provide semantically correct results ambiguity in the user-provided natural language queries must be 
adequately addressed. This thesis investigates various approaches to ambiguity management and
resolution. Natural language queries can demonstrate ambiguity even at low levels of complexity —
e.g. there are two different types of "sales" in a database schema, and the user asks to retireve
"all sales".

Such situations present the second major challenge associated with the practical implementation of natural
language database interfaces. The success of this concept will significantly depend on whether suitable 
designs and mitigation techniques can be implemented without creating problems with regards to the 
aforementioned performance and hardware requirements. The research focus lies on both preventative measures
through optimized pre-processing stages and prompt engineering techniques as well as reactive strategies
that post process LLM output, either on the basis of further user input or context inference.

\subsubsection*{RQ3 — Which strategies are increasing semantic accuracy of queries?}

In order to enhance the semantic accuracy a series of improvements may be applied to the pipeline.
Potential optimizations include supplying (parts of) the schema during LLM prompting, implementation of
interactive contextual reasoning through a conversational interface which would allow for user
refinement, the implementation of a robust SQL parsing and validation mechanism and a hybrid approach
partly relying on traditional NLP preprocessing techniques. This research will quantify semantic accuracy
using popular NL2SQL benchmarks and empirically evaluate the impact each approach has on the benchmark
performance. Furthermore this research will take a look at the optimal combination of the aforementioned
solutions in order to develop a system that strikes the right balance between accuracy and performance.

% \subsubsection{Secondary Research Questions}

% \subsubsection*{RQ4 — What use cases are most suitable for natural language database interfaces?}

% \subsubsection*{RQ5 — Where are the limitations of natural language database interfaces?}

% \subsubsection*{RQ6 — How is NL2SQL model performance impacted through model selection, fine tuning and pre/post processing?}

\newpage

\subsection{Structure of the Thesis}

This thesis is following a research and development methodology in order to implement a natural language
interface for databases, in particular postgres is used.

\begin{enumerate}
    \item \textbf{Literature Review} — An analysis of the existing research in the fields of
          natural language interfaces (NLI) for databases, GPU integration for acceleration
          of database operations, and LLM/AI Model integration within database systems.
          This phase establishes the theoretical foundation for this research and identifies current 
          state-of-the-art approaches, their benefits and shortcomings.
     \item \textbf{Decomposition \& Requirements} — Decomposing the problem statement into its
          fundamentals and deriving system requirements for the design phase from it. The goal
          of this section is to arrive at a list of functional and non-functional requirements that
          must be taken into account and fulfilled by the design and implementation phases respectively.
     \item \textbf{System Design} — Design of a system architecture that can utilize GPU acceleration
          for LLM integration from within postgres. The primary goals of the system design phase
          are to arrive at an architecture that yields low latency natural language processing,
          schema-aware SQL query generation, ambiguity detection and resolution whilst maintaining
          a high semantic accuracy.
     \item \textbf{Implementation} — The implementation of a PostgreSQL extension according to the
          above system design that relies on \texttt{rust} and \texttt{pgrx}. This extension will
          provide a GPU accelerated framework for executing LLMs, implement a natural language
          to query generation pipeline that relies on the SQL schema and create database functions
          and operators for both query generation and execution.
     \item \textbf{Evaluation and Benchmarking} — An assesment framework and benchmark that introspects
          the implementations performance in multiple dimensions. Namely the most relevant dimensions
          for this thesis are:
          \begin{enumerate}
              \item Semantic Accuracy — Measuring the overall accuracy of results delivered for a given
                    natural language input.
              \item Ambiguity Resolution Capabilities — How well the system performs when confronted with
                    ambiguous natural language input and database schemas.
              \item Performance Metrics — Measuring the latency, throughput and resource utilization 
                    of the implementation.
          \end{enumerate}
      \item \textbf{Discussion} — Analysis and interpretation of the evaluation phase results against
            the research goals of this thesis. Evaluating the performance and accuracy results recorded
            during the benchmarks against the question whether real world deployments of NILs are feasible.
            Furthermore the effectiveness of ambiguity resolution capabilities and semantic accuracy enhancement
            strategies are showing a statistically significant effect.
       \item \textbf{Summary and Outlook} — Summarizes the contributions, addresses limitations
            of this thesis and the implementation, and proposes directions for future research alongside
            possible applications. Primary future research topics include advanced GPU optimization
            techniques (e.g. further quantization), accuracy and performance impact of model fine tuning,
            techniques, scalability of such a system in enterprise scenarios and the evaluation of security
            and privacy considerations (e.g. managing access control).
\end{enumerate}

\newpage

% Theoretical Foundations
\section{Literature Review}

In this section a comprehensive literature review is performed to asses the research landscape on NL2SQL
(sometimes also referred to as Text-to-SQL or T2SQL) and NLIDBs. From the time their development accelerated in
the late 1990s and early 2000s \citep{NLIDBs, NLIDBTheory, ILPParsing, ILPParsing2} until now, observing multiple 
larger paradigm shifts happening over time \citep{GRAPPA, STRUG, Seq2SQL, NALIR, SQLizer}. In particular this 
research focuses on the recent advancements when it comes to language models and how they can be harnessed for
effective NL2SQL systems \citep{LLM-Sql, T2SQL-LLM-Bench, T2SQL-LLM-Bench-2, T2SQL-LLM-Bench-3, SPIDER2, BIRD}.

This literature review is covering the foundational concepts, challenges, key advancements and research gaps
associated with using natural language instead of SQL. It lays the foundation for this thesis and helps to set
the research questions introduced in the previous chapter in context.

% bessere einleitung

\subsection{Foundations of Natural Language Interfaces to Databases}

Earlier papers in the research landscape on Natural Language Database Interfaces (NLIDBs) date over half
a century back, into the early 1970s. Two decades after the first major research systems where developed
in this domain, \citeauthor*{NLIDBs} have published an introduction and an overview over NLIDBs where an overview of
state-of-the-art approaches were provided. \citep{NLIDBs} Their work outlined multiple key issues and challenges
associated with NLIDBs, and compared them against existing / competing solutions like formal query languages,
form-based interfaces and graphical interfaces. These challenges (like unobvious limits, linguistic ambiguities,
semantic inaccuracy, tedious configuration etc.) have shaped this field of research and are still considered relevant
metrics today.

Early NLIDBs primarily relied on traditional natural language processing (NLP) techniques in order to achieve
natural language understanding capabilities. With \textsc{Chill} an inductive logic programming (ILP) approach
was first introduced for NL2SQL systems, marking one of the key events when it comes to machine learning usage.
\citep{ILPParsing} In \citeyear{ILPParsing2} \citeauthor*{ILPParsing2} have extended the approach of ILP
parsing for natural language database queries with multi clause construction, yielding promising results
in the field of NLIDBs. \citep{ILPParsing2}

Building on the systematic overview of \citeauthor*{NLIDBs} and the first machine learning approaches from
\citeauthor*{ILPParsing} aswell as \citeauthor*{ILPParsing2}, \citeauthor{NLIDBTheory} have proposed a novel 
approach for implementing NLIDBs and outperformed at the time state-of-the-art solutions from \cite{ILPParsing} 
\cite{ILPParsing2} — achieving 80\% semantic accuracy. \citep{NLIDBTheory} The novelty of the \textsc{Percise}
system lies in its natural language processing approach, specifically its lexical mapping strategy, allowing 
\textsc{Percise} to identify questions it can, and can't answer (introducing the concept of \textit{semantically 
tractable questions}) which therefore results in a better and interactive end user experience. Their experiements 
also showed that this approach is \textit{transferrable} and \textit{unbiased} — it is possible to feed in new, 
unknown questions into the system and maintain performance characteristics, whereas it was shown that 
\cite{ILPParsing} were suffering from a distribution drift of the questions asked. \citep{NLIDBTheory}

The theoretical foundations and research questions highlighted by the aforementioned works, shaped the research
field and highlighted the following, ongoing research:

\begin{enumerate}
    \item The trade-off characteristics derived from choosing a machine learning vs. traditional NLP approach (e.g. 
          \textsc{Chill} versus \textsc{Percise}). E.g. coverage versus correctness. \citep{ILPParsing, NLIDBTheory}
    \item The linguistic challenges associated with bringing NLIDBs into use (e.g. semantic inaccuracy, linguistic 
          ambiguity, unclear language coverage etc.) \citep{NLIDBs}
    \item The value of systems and approaches which double down on reliability and semantic accuracy rather than giving
          promising but incorrect answers. \citep{NLIDBs, NLIDBTheory}
\end{enumerate}

Fundamentally this highlights the tension and mismatch between the characteristics of natural language, which is 
able to be ambiguous, \textit{semantically untractable} or able to be incomplete in meaning and formal languages
like SQL which always have on deterministic and \textit{semantically tractable} meaning they convey in each statement. 
As Schneiderman and Norman have pointed out according to \citeauthor*{NLIDBTheory}, users are ``unwilling to trade 
reliable and predictable user interfaces for intelligent but unreliable ones'' which induces performance expectations
on NLIDB implementations to be highly certain about the questions it can, and can't answer, whilst maintaing as high
as possible natural language coverage. \citep{NLIDBTheory}

\subsection{Traditional NL2SQL Approaches}

Prior to the wide-spread dominance of machine learning approaches for natural language processing a variety of
traditional, rather discrtete approaches have been explored in the field of NL2SQL / NLIDBs. These discrete
approaches have laid the foundations for transitioning towards the application of machine learning techniques
for NL2SQL.

\subsubsection{Rule-based and Grammar-based Systems}

Foundational research of NL2SQL system mostly focused around applying rule engines that were tedious to set up and
expensive to maintain / transfer across database systems. These rule engines mostly relied on the systematic 
identification of linguistic patterns / were trying to template SQL from information that was derived from processing the 
natural language query. \citep{Rendezvous, Lunar, Ladder} These approaches mostly tried to formalize natural language 
queries into formal grammers which could then be deterministically mapped into a valid SQL query. \citep{Lunar} These 
approaches have strong downsides when it comes to the variety of natural language constructs they can process, aswell as 
runtime adoption of new / unknown databases, query constructs etc. A potential upside of this class of NL2SQL systems is 
that they can confidently and reproducably identify questions they can, and can't answer — thus leading to very reliable 
and predictable user interfaces.

\subsubsection{Semantic Parsing using String-Kernels}

A significiant milestone in parsing techniques of natural language queries was reached by \citeauthor*{StringKernels} in 
\citeyear{StringKernels}. The introduction of string kernels for semantic parsing represented a novel achievement, when 
it comes to fusing discrete parsing approaches using a formal grammer like LSNLIS developed by \cite{Lunar} and 
learning / training approaches to understand unseen language patterns / unknown natural language query structures. This 
allowed for more flexible pattern recognition when compared to traditional rule-based systems.

The core innovative characteristic of this approach lies in its capability to understand similarities between natural 
language expressions based on subsequence patterns rather than relying on exact matches. This made \textsc{Krisp}, the 
research NLIDB system developed by \cite{StringKernels} much more robust to language variations in phrasing and noise 
(e.g. spelling mistakes) in the input. As the \citeauthor*{StringKernels} demonstrated through experiments on real-world 
datasets, this approach compared favorably to existing systems of the time like \textsc{Chill}, especially in handling 
noisy inputs — a frequent challenge rigid rule-based systems faced in real world scenarios \citep{StringKernels, 
ILPParsing}.

\subsubsection{Graph Matching Methods}

\cite{GraphMatching} brought together several research threads and reapplied emerging graph matching research models
to natural language processing, specifically to natural language queries. Graph matching was applied once the natural 
language query was parsed using a Combinatory Categorial Grammar (CCG) approach into a semantic graph which denotes the 
relationship between semantic entities in it. This graph could then be matched against the actual graph derived from 
the database, since they share topological traits that can be used for matching \citep{GraphMatching}. This approach 
allowed to apply querying systems without having any question-answer pairs or manual annotations for training the system, 
which implies easier scalability / transferability across domains, since the system does not require any additional 
tweaks.

Even though this approach was novel and showed improved the performance over existing state-of-the-art approaches, it
was showing that graph matching quickly reaches its limitations. This approach relied heavily on the CCG parser's 
accuracy, with parsing errors accounting for 10-25\% (depending on the dataset) of system failures \citep{GraphMatching}. 
Furthermore it struggled with both ambiguous language constructs and potential mismatches between natural language 
representation of relationships and database layouts — more complicated database designs, which may not match the users 
intuitive understanding resulted in a different topology and hence could not be matched \citep[p.~387]{GraphMatching}.

\subsubsection{Interactive Systems}

In \citeyear{NALIR} \citeauthor*{NALIR} identified that perfect translation of natural language into SQL was challenging
due to natural language not being made for query expressions as it heavily relies on contextual information and clarifying
questions in order to disambiguate conversations \citep{NALIR}. These learnings relate to early prior art from 
\citeauthor{UnnaturalQueryLanguage} and \citeauthor{Rendezvous} which also made this observation — ``natural language
is not a natural query language.'' \citep{UnnaturalQueryLanguage}. The solution introduced by NaLIR further emphasized
how important an interactive, conversational usage model is, when offering a natural language interface \citep{NALIR}.

NaLIR could accept logically complex English language sentences as input and translate them into SQL queries with various 
complexities, including aggregation, nesting, and different types of joins etc. The key innovative characteristic of 
NaLIR lies in its interactive communication mechanism (much like \textsc{Rendezvous}) that could detect potential 
misinterpretations and engage users to resolve ambiguities present in their natural language query without forcing them
to entirely rephrase their query \cite{NALIR}. This approach, while showing awareness for its limitations (with regards
to entirely automating / deriving SQL generation from potentially ambiguous or faulty user input) showed that it was 
possible to overcome these limtations through choosing the right interaction model — ``In our system, we generate
multiple possible interpretations for a natural language query and translate all of them in natural language for the
user to choose from'' —, rather than optimizing the generation part of the system \cite{NALIR}.

\subsubsection{Query Synthesis}

\cite{SQLizer} introduced SQLizer, which synthesizes SQL queries from natural language \citep{SQLizer}. This paper
presents a novel approach when it comes to NL2SQL as it is merging prevalent semantic parsing techniques (outlined above)
with an program synthesis (or query synthesis) approach. SQLizer makes use of a three stage processing model for 
natural language models: first generating a sketch of the query using semantic parsing, then using type-directed
synthesis to complete the sketch and finally using automated repair, if required.

\citeauthor*{SQLizer} show that alternating between repairing and synthesis yields results that beat state-of-the-art
NL2SQL approaches like NaLIR. SQLizer is fully automated and database-agnostic, requiring no knowledge of the underlying
schema. The authors evaluated SQLizer on 455 queries across three databases, where it ranked the correct query in the
top 5 results for roughly 90\% of the queries. This represents a significant improvement over NaLIR \citep{NALIR},
the previous state-of-the-art system \citep{SQLizer}.

Potential short commings of this approch include queries which yield empty results, dealing with language variations
as SQLizer is still using semantic parsing, and domain-specific terminology, all while still requiring users
to select from multiple query options which reduces the overall usability of the system \citep[p.22-23]{SQLizer}.

\subsubsection{Limitations of Traditional Approaches to NL2SQL}

Despite being innovative and achieving state-of-the-art results, many of the above outlined approaches face severe 
challenges when moving outside of an research environment. Many of these systems performed comparatively good on research
benchmarks that were often composed of controlled question types and limited data variety. Ultimatively no standard
benchmark existed for NL2SQL in this era, hence comparing different NL2SQL systems against each other is a problem on 
its own. Despite not having a standard benchmark that all approaches could be unifiably evaluated against, several 
fundamental challenges emerged / remained with these approaches:

\begin{enumerate}
    \item \textbf{Limited linguistic coverage} — Prevalent rule-based and semantic-parsing based systems were only able to 
          process the a small subset of the natural language they were programmed for. This severely limited their 
          ability to handle different phrasings of the same end-user goal \citep{StringKernels, UnnaturalQueryLanguage, 
          Lunar, Ladder}.
    \item \textbf{Transferability} — Traditional approaches typically required extensive manual configuration or at least
          a training phase / adaption for each database they were deployed for, hindering cross domain usage through being 
          expensive and time-consuming to adapt \citep{NLIDBs, Lunar}.
    \item \textbf{Brittleness} — Many of the systems introduced in this subchapter did not handle synonyms, paraphrasing,   
          or spelling errors well. Manual adaption / handling was needed in order to becomes resilient against each class
          of problems \citep{StringKernels, SQLizer}.
    \item \textbf{Poor scalability} — With potentially more complex underlying databases, traditional 
          solutions often showed to perform worse. \citeauthor*{GraphMatching} found, that with increasing schema
          complexity more compute was required to resolve the natural language query to a suitable query candidate
          making them less transferable and scalable than initially anticipated \citep{GraphMatching} — ``Evaluating on
          all domains in Freebase would generate a very large number of queries for which denotations would have to be 
          computed ... Our system loads Freebase using Virtuoso and queries it with SPARQL. Virtuoso is slow in dealing with
          millions of queries indexed on the entire Freebase, and is the only reason we did not work with the complete
          Freebase.'' which indicates underlying system design issues with runtime complexity.
\end{enumerate}

These flaws of traditional NL2SQL approaches made it apparent, that a different class of approaches is needed, which 
increase transferability and reduce the brittleness since users are ``unwilling to trade reliable and predictable user 
interfaces for intelligent but unreliable ones'' according to \cite{NLIDBTheory}. Whilst many approaches outlined tractable 
ways to increase user satisfaction and accuracy (like \citeauthor*{Rendezvous} did in \citeyear{Rendezvous} with a 
conversational approach), NLIDBs were and are not considered to be a solved problem.

\subsection{Neural NL2SQL Approaches}

\subsection{Benchmark Evolution}

\subsection{LLM-based NL2SQL Systems}

\subsection{Optimization for LLM-based NL2SQL}

\subsection{Complex Techniques in NL2SQL}

\subsection{Research Gaps}

\newpage

\section{Decomposition \& Requirements}

\subsection{Problem Decomposition}

\subsection{Requirements}

% Conceptual Design
\section{System Design}

\subsection{Architecture Design}
\subsubsection{Interface Design}
\subsubsection{Data Model}
\subsubsection{Integration into SQL}

\subsection{Technical Implementation Strategies}

\newpage

% Implementation
\section{Implementation}

\subsection{Development Environment and Tools}

\subsection{Integration of the Model}

\subsection{Development of the PostgreSQL Extension}

\subsection{Optimization}

\newpage

% Evaluation
\section{Evaluation}

% https://github.com/petavue/NL2SQL-Benchmark/blob/main/results/Report/NL2SQL%20Benchmark%20Report.pdf

\subsection{Test Environment and Methodology}

\subsection{Performance Tests}
\subsubsection{Latency}
\subsubsection{Throughput}
\subsubsection{Scalability}

\subsection{Use Cases}
\subsubsection{Natural Language Queries}
\subsubsection{Text Generation Within the Database}
\subsubsection{Semantic Search and Text Classification}

\subsection{Comparison with Alternative Approaches}

\newpage

% Discussion
\section{Discussion}

\subsection{Interpretation of Results}
\subsection{Limitations of the Implementation}
\subsection{Ethical and Data Privacy Considerations}
\subsection{Potential Future Developments}

\newpage

% Summary and Outlook
\section{Summary and Outlook}

\subsection{Summary of Results}
\subsection{Addressing the Research Questions}
\subsection{Outlook for Future Research and Development}

\newpage

\newpage

\appendix
\section*{Appendix}
\subsection*{Installation Guide}
\subsection*{API Documentation}
\subsection*{Code Examples}
\subsection*{Test Data and Results}

\nocite{*}

\bibliographystyle{apacite}
\bibliography{references}

\end{document}