### 2025-10-01 thesis/chapters/02-literature-review/index.tex

\section{Literature Review}

%\input{chapters/02-literature-review/foundations/nl-interfaces}
%\input{chapters/02-literature-review/traditional-approaches/traditional-approaches}
%\input{chapters/02-literature-review/neural-approaches/neural-approaches}
%\input{chapters/02-literature-review/pretrained-models/pretrained-models}
%\input{chapters/02-literature-review/llm-approaches/llm-approaches}
%\input{chapters/02-literature-review/benchmarking/benchmarking}
%\input{chapters/02-literature-review/research-gaps/research-gaps}

% https://github.com/HKUSTDial/NL2SQL_Handbook/blob/main/assets/river.svg

In this section a comprehensive literature review is performed to asses the research landscape on NL2SQL
(sometimes also referred to as Text-to-SQL or T2SQL) and NLIDBs. From the time their development accelerated in
the late 1990s and early 2000s \citep{NLIDBs, NLIDBTheory, ILPParsing, ILPParsing2} until now, observing multiple 
larger paradigm shifts happening over time \citep{GRAPPA, STRUG, Seq2SQL, NALIR, SQLizer}. In particular this 
research focuses on the recent advancements when it comes to language models and how they can be harnessed for
effective NL2SQL systems \citep{LLM-Sql, DAIL-SQL, T2SQL-LLM-Bench-2, T2SQL-LLM-Bench-3, Spider2, BIRD}.

This literature review is covering the foundational concepts, challenges, key advancements and research gaps
associated with using natural language instead of SQL. It lays the foundation for this thesis and helps to set
the research questions introduced in the previous chapter in context.

\subimport{}{foundations}
\subimport{}{traditional-approaches}
\subimport{}{neural-approaches}
\subimport{}{pretrained-models}
\subimport{}{llm-approaches}
\subimport{}{benchmarking}
\subimport{}{research-gaps}