### 2025-10-01 thesis/chapters/03-theoretical-foundations/index.tex

\section{Theoretical Foundations}

This section serves as an introduction to the theoretical concepts utilized in this thesis. It is meant to establish a
baseline of understanding for the design, implementation and evaluation phases. It briefly covers relational database theory
(incl. SQL), machine learning fundamentals and their subsequent application for large language models.

\subimport{}{relational-theory.tex}
\subimport{}{machine-learning.tex}
\subimport{}{llms.tex}

%\subsection{Problem Decomposition}

%\subsection{Requirements}

% Conceptual Design

%\subsection{System Design}

%\subsubsection{Architecture Design}

%\subsection{Technical Implementation Strategies}