### 2025-10-01 thesis/literature-review/neural-approaches.tex

\subsection{Neural NL2SQL Approaches}

The previously outlined limitations of traditional approaches to solving NL2SQL / implementing NLIDBs pushed the research
branch around neural network application forward to step in and propose new solutions which address the brittleness, 
transferability and scalability concerns addressed with logical programming approaches. Neural approaches showed to yield
significant improvements in terms of transferability and overall accuracy which led to a paradigm shift in this research
field.

\subsubsection{Early Neural Approaches}

In \citeyear{Seq2SQL} \citeauthor*{Seq2SQL} released Seq2SQL which represents a significant breakthrough and leap in NLIDB 
research. Seq2SQL was an early research system that in the field of neural network application and as one of the first papers
to frame the implementation of NLIDBs / NL2SQL Systems as a reinforement learning problem. The system utilized iterative query 
execution in the reward function to improve its accuracy \citep{Seq2SQL}. In the same paper \citeauthor*{Seq2SQL} introduced 
WikiSQL, a training dataset, which enables large scale (in \citeyear{Seq2SQL}) model training.

SQLNet \citep{SQLNet} addressed primarily the order-sensitivitiy trait of Seq2SQL \citep{Seq2SQL} that was prevalent due to being
a derivative approach from sequence-to-sequence approaches. SQLNet diverges from sequence-to-sequence and joins multiple research 
threads, employing a sketch-based query generation. SQLNet breaks down complex queries into smaller (hence more manageable)
sub-queries which can then be individually sketched and refined, yielding a system that outperformed state-of-the-art by 9\% 
to 13\% \citep{SQLNet}.

\citeauthor*{TypeSQL} have introduced \textsc{TypeSQL}, a variation of the SQLNet-approach, in \citeyear{TypeSQL}.
\textsc{TypeSQL}'s primary difference to SQLNet is the encoding of type information for SQL generation. The approach scanned
for entity references and values in natural language and was able to improve performance by 5.5\% over SOTA-Models like
SQLNet whilst requiring significantly less training time, indicating that type information was a useful information for deriving 
accurate SQL queries from user input \citep{TypeSQL}.

\subsubsection{Intermediate Neural Developments}

Later in \citeyear{SyntaxSQLNet} \citeauthor*{SyntaxSQLNet} released SyntaxSQLNet, a followup research to \textsc{TypeSQL},
which represented a slight change in approach and research focus. In direct comparison SyntaxSQLNet focused primarily around
complex query generation using a syntax tree decoder, allowing for longer and more cohesive query generation \citep{SyntaxSQLNet}.
This advancement over \textsc{TypeSQL} allowed more complex queries to be reliably generated, enabling multiple clauses aswell as
nested queries. SyntaxSQLNet was one of the earlier research efforts which utilized Spider instead of WikiSQL (introduced by 
\cite{Seq2SQL}), a large-scale NL2SQL dataset, incorperating 10.181 hand annotated natural language question and alongside
5.693 unique SQL examples that spread across 138 different domains \citep{Spider}. This research led the transition of
comparatively simple, research-grade, neural systems for NLIDBs towards systems which are feasible in the real world.

Building on the above approaches, \citeauthor*{IRNet} have introduced IRNet, a neural network approach using intermediate
representation as a bridge between natural language and SQL in which semantic queries could be expressed. The intermediate
format SemQL (or semantic query language) was utilized to transform and synthesize queries on the actual database schema more
accurately than Seq2SQL. IRNet followed a three phase approach: schema linking between the natural language query and database
layout, synthesis of SemQL as intermediate representation and deterministic conversion of SemQL to SQL. This approach allowed
IRNet to outperform state-of-the-art approaches on the \textsc{Spider} benchmark by 19.5\%, placing IRNet at an overall accuracy
of 46.7\% \citep{IRNet}.

Following IRNet, graph neural networks (GNN) have been explored as alternative architecture by \cite{GNN}, representing the
database schema as a graph and using message passing to model relationships between tables, columns and natural language input.
This approach demonstrated the capability to improve reasoning and query generation capability. \citeauthor{GNN} showed that when 
evaluating against the \textsc{Spider} benchmark GNN outperforms both SyntaxSQLNet (and therefore \textsc{SQLNet} and
\textsc{TypeSQL}). Although presenting a signficiant advancement over previous state-of-the-art approaches, GNN falls behind in
performance against IRNet by 6\% \citep{IRNet, GNN}.

\subsubsection{Relation-Aware Transformer Approaches}

The release of RAT-SQL (Relation-Aware Transformer for SQL) \cite{RATSQL} represents the most significant leap in research of
neural NL2SQL approaches. RAT-SQL diverged from earlier research through emphasizing the relationship between natural language
and the database schema elements using relation-aware self-attention representing a novel approach for solving \textit{schema 
linking} \citep{RATSQL}.

RAT-SQL's primary innovations was the ability to infer, understand and utilize the relationship between individual tokens in the 
natural language query and link it to the database schema. Thus allowing for reasoning capabilities on the actual database schema
while generating the query.

This architecture yielded a 57.2\% in exact match accuracy when being evaluated on the \textsc{Spider} benchmark, substantially
outperforming comparative approaches like GNN, IRNet and IRNet V2 by 10.5\%, 9.8\% and 8.7\% respectively. Although overall
accuracy improved across all approaches when being paired with BERT (Bidirectional Encoder Representations from Transformers, a 
popular pre-trained lanugage model from Google) the $\delta$ between the indivudual approaches remained relatively steady, leaving
RAT-SQL outperforming state-of-the-art approaches by 5\% to 12.2\% further demonstrating the capability advancement yielded by 
this system \citep{RATSQL}.

\subsubsection{Comparative Analysis of Neural Approaches}

The evolution from early neural approaches to RAT-SQL emphasized the rapid advancements that happened in the research field of 
neural NL2SQL approaches in different dimensions:

\begin{enumerate}
    \item \textbf{Model Complexity} — Given the research progression from early sequence to sequence translation approaches 
          (\textsc{Seq2SQL}) towards sketch based and type augumented and graph based approaches (\textsc{TypeSQL, SQLNet, GNN})
          and syntax tree decoding emphasized by \textsc{SyntaxSQLNet}, neural approaches continously advanced in the complexity
          of approaches that is required to beat state-of-the-art approaches in contemprorary benchmarks like \textsc{Spider}. RAT-
          SQL presents one of the late and most complex advancements in the field of neural NL2SQL approaches with its adapted 
          self attention mechanism \citep{Seq2SQL, TypeSQL, GNN, SyntaxSQLNet, RATSQL}.
    \item \textbf{Transferability} — Each of the approaches introduced above represents a succession in terms of their 
          transferability. The field of neural NL2SQL approaches significantly improved the ability for NLIDBs to generalize over 
          the underlying database schemas. RAT-SQL showed the strongest cross-domain accuracy (that is benchmarked by the 
          \textsc{Spider} benchmark). With standard benchmarks emerging it became easier to verify and quantify which approach had 
          the highest transferability as \textsc{Spider} specifically had independent development and test datasets, preventing 
          approaches from over-optimizing on training data \citep{, RATSQL}.
    \item \textbf{Robustness} — As research systems advanced in complexity and shifted from raw input to output translation
          (Seq2SQL) their robustness steadily increased. Through more approaches like \textsc{SyntaxSQLNet} which utilized 
          structured decoding, \textsc{IRNet} which relied on an intermediate representation and \textsc{RAT-SQL} the challenges 
          around \textit{schema linking} outlined by \cite{RATSQL} have increasilingly led to more robust systems that can handle 
          rephrasings, spelling mistakes and variations in natural language usage far beyond what traditional NL2SQL approaches 
          could accomplish \cite{SyntaxSQLNet, IRNet, RATSQL}.
    \item \textbf{Query Complexity} — The performance on complex queries involving multiple tables, relying on complex 
          aggregations, nested structures and joins dramatically improved over the course of the research that happened in this 
          field. Whilst \textsc{IRNet} reresents one of the first singificant advancements when it comes to the ability of neural 
          approaches to handle complex queries, \textsc{RAT-SQL} still showed to outperform the intermediate representation 
          approach introduced by \textsc{IRNet} by up to 10.5\% \citep{IRNet, RATSQL}.
    \item \textbf{Schema Understanding} — Whilst early approaches like Seq2SQL primarily applied reinforcement learning for end to 
          end query generation \citep{Seq2SQL}, later approaches like \textsc{TypeSQL, GNN} and specifically \textsc{RAT-SQL} 
          showed novel and state-of-the-art \textit{schema understanding / schema linking} capabilities, yielding the ability to 
          accurately reason about user intent and traverse the database schema while generating queries \citep{TypeSQL, GNN, 
          RATSQL}. 
\end{enumerate}

\subsubsection{Limitations of Neural Approaches}

Despite the dramatic \textit{accuracy, transferability} and \textit{robustness} improvements that could be observed with late
neural approaches \citep{IRNet, RATSQL}, neural approaches still suffered from serious shortcommings / unsolved challenges:

\begin{enumerate}
    \item \textbf{Training Data} — Utilizing neural networks these approaches required susbtantially more training data 
          (ie. natural language paired with output SQL queries) than traditional systems which required serious efforts of data 
          collection \citep{Spider}.
    \item \textbf{Correctness} — The inherent mismatch between neural networks and formal languages yielded cases where models 
          produced invalid SQL code. Approaches like \textsc{SyntaxSQLNet} improved the tried to solve this circumstance by 
          utilizing syntax trees during decoding but nonetheless syntactic correctness remained a challenge across future
          iterations of neural systems. \citep{SyntaxSQLNet}
    \item \textbf{Domain Language} — Despite increased \textit{transferability} characteristics neural approaches still suffered
          from a limited vocabulary and inter-domain understanding of terminology and relation between concepts which made highly 
          domain specific natural language queries challenging.
    \item \textbf{Observability} — The black-box nature of neural networks made approaches relying on them, particularly the ones 
          with complex architectures, hard do understand / explain in case when neural systems yielded undesirable output. 
\end{enumerate}

The introduction and advancement of early neural NL2SQL approaches led to significant advancements in the research and feasibility
of NLIDBs. The research shift started in this era established the foundations for further and more advanced machine learning 
approaches (specifically language model oriented approaches )being researched. Neural approaches showed to significant 
improvements in performance when being paired with pre-trained language models \citep{RATSQL} which led to further research on 
their applicability.