### 2025-10-01 thesis/theoretical-foundations/machine-learning.tex

\subsection{Machine Learning Fundamentals}

Machine learning represents the algorithmic foundation for state-of-the-art NL2SQL systems, enabling systems to learn how natural
language queries translate into SQL. Understanding these foundations is essential for implementing and optimizing large language
model-based NL2SQL approaches.

\subsubsection{Neural Network Architecture}

Neural networks are the computational concept behind contemporary NL2SQL approaches, representing complex capabilities through
compositions of simpler operations. A neural network consists of interconnected computational units (called neurons) organized
in layers, where each connection has an associated (and learnable) weight and bias parameters.

The fundamental computation paradigm of neural network outputs is called forward propagation, which is applying the respective weights ($W$)
and biases ($b$) to the input parameter ($a$) and transforming it using a (non-linear) activation function $f$:

\begin{equation}
a^{l+1} = f(Wa^l + b)
\end{equation}

where $W \in \mathbb{R}^{m \times n}$ is the weight matrix, $a \in \mathbb{R}^{n}$ is the input vector, $b \in \mathbb{R}^{m}$ is
the bias vector, and $f$ is an activiation function (e.g. ReLU, sigmoid, or tanh). Using this function repeatedly, propagates information
through the neural network.

The most important architectural components for language processing models include:

\begin{itemize}
    \item \textbf{Feedforward Networks} — Which can process fixed-size inputs through successive linear transformations and activations,
          this architecture is particularly suitable for classification and regression problems within NL2SQL systems.
    \item \textbf{Recurrent Networks (RNNs/LSTMs)} — Which can handle sequential data of variable-length by maintaining hidden states
          that capture dependencies, enabling processing of natural language sequences of arbitrary length.
    \item \textbf{Embedding Layers} — Which map discrete tokens (e.g. words, characters, or subwords) to a dense vector representation,
          thus providing the foundation for subsequent neural language processing.
\end{itemize}

The ability of neural networks to learn hierarchical representations through multiple layers makes them particularly well-suited for
the complex translation required in NL2SQL systems.

\subsubsection{Sequence-to-Sequence Models}

Sequence-to-sequence (Seq2Seq) models established the foundational architecture for neural NL2SQL approaches \citep{Seq2SQL, SQLNet}.
These models which rely on RNNs and LSTMs addressed the fundamental challenge of mapping variable-length input sequences (natural
language queries) to variable-length output sequences (SQL statements).

The Seq2Seq model architecture is composed out of three essential components:

\begin{enumerate}
    \item \textbf{Encoder} — Process the input natural language sequence $x_1, x_2, \ldots, x_n$ into a sequence of hidden
          representations $h_1, h_2, \ldots, h_n$, capturing the semantic content of the query.
    \item \textbf{Decoder} — Generates the output SQL sequence $y_1, y_2, \ldots, y_m$ autoregressively, where each token is
          predicted based on previous outputs and encoder representations.
    \item \textbf{Attention} — Allows the decoder to selectively focus on relevant portions of the input sequence, addressing
          the information bottleneck of fixed-size context vectors.
\end{enumerate}