### 2025-10-01 thesis/theoretical-foundations/relational-theory.tex

\subsection{Relational Database Theory}

The relational database theory provides a mathematical framework for formalizing data layout in database management
systems (DBMS), enabling efficient storage and retrieval mechanisms. This theory is helps to understand the way in
which natural language queries need to be translated into database operations and what challenges arise during this
translation.

\subsubsection{Relational Model Fundamentals}

The relational model was introduced by \citeauthor{RelationalModel} in \citeyear{RelationalModel} and essentially forms
the theoretical foundation for most modern database systems \citep{RelationalModel, Rendezvous}. A relational database
organizes data in relations (ie. tables), where each relation consists of tuples (ie. rows) containing attributes
(ie. columns) of specific domains (ie. data types and constraints). The mathematical origins of this model root in
set theory and first-order predicate logic \citep{RelationalModel, DatabaseSystemIntro}.

Furthermore the relational model abstracts data storage and respective physical implementation details by design,
allowing interactions with data through respective logical structures (relations, tuples, attributes etc.) rather than
exposing the underlying storage mechanism. This abstraction is fundamental for NL2SQL systems, as natural language
queries can be mapped into queries on these structures regardless of underlying physical storage implementation,
allowing for better transferability of NL2SQL systems (e.g. across different database deployments).

\subsubsection{Core Concepts}

The core concepts of the relational model that are relevant for this thesis include:

\begin{itemize}
    \item \textbf{Relations} — Mathematical sets of tuples representing entities (e.g. employees, products) or
          relationships (e.g. enrollment, purchases). Each relation has a fixed structure.
    \item \textbf{Tuples} — Tuples represent rows of data, by grouping a several attributes into one logical unit
          (e.g. one specific purchase, containing the amount, date and shop). All tuples in the same relation share
          the same structure.
    \item \textbf{Attributes} — Attributes are frequently referred to as columns, representing a specific dimension
          of a tuple (e.g. customers have names). Attributes are represented in a fixed domain.
    \item \textbf{Domains} — Domains are sets of allowed values for attributes, including mathematically formalizing
          data types and constraints.
    \item \textbf{Schemas} — Structural definitions specifying relations, their attribute names, domains, and integrity
          constraints.
    \item \textbf{Primary Keys} — Attributes that uniquely identify tuples within a relation. These
          ensure entity integrity and ensure referential relationships.
    \item \textbf{Foreign Keys} — Attributes referencing primary keys in other relations, allowing databases to maintain
          referential integrity and enabling semantically correct complex queries across multiple tables through joins.
\end{itemize}

These concepts form the semantic foundation that natural language interfaces must navigate when translating user queries
into formal database operations. As shown in the literature review \textit{schema-linking} is a crucial problem of
NLIDBs — establishing an understanding of which relations a natural language query refers to is the first step of
formalizing a database operation for retrieval.

\subsubsection{SQL as a Declarative Query Language}

Structured Query Language (SQL) is a defacto standard for interfacing with relational databases. It is a declarative
language to describe database operations like selection, insertion, updating and deletion of database conte/home/mara/Workspace/thb/thesis/main.texnts. SQL's
design is strongly influenced by relational algebra and tuple relational calculus. It enables complex data retrieval
through a readable query syntax incorporating relational algebra operations like:

\begin{itemize}
    \item \textbf{Selection} ($\sigma$) — Filtering tuples based on specified conditions
    \item \textbf{Projection} ($\pi$) — Extracting specific attributes from relations
    \item \textbf{Joins} ($\bowtie$) — Combining data from multiple relations based on common attributes
    \item \textbf{Aggregation} — Computing summary statistics over groups of tuples
\end{itemize}

The basic syntax of SQL queries follows a logical pattern that reflects these theoretical operations:

\begin{verbatim}
SELECT attributes      -- Project attributes
FROM relations         -- Specify relations to use
WHERE conditions       -- Selection of tuples
GROUP BY attributes    -- Grouping for aggregation
HAVING conditions      -- Group-level filtering
ORDER BY attributes    -- Ordering of results
\end{verbatim}

There are further operations like various types of \texttt{JOIN}s, \texttt{OFFSET}, \texttt{LIMIT}, \texttt{WITH}
that are heavily used in SQL which are excluded for readability and simplicity.

The nature of SQL creates a fundamental challenge for NL2SQL systems; natural language queries express (often fuzzy) user
intent in terms of desired output (e.g., ``Give me 5 great movies'') while SQL formalizes an explicit and discrete way
of how to retrieve this information from the database:

\begin{verbatim}
SELECT m.title, AVG(r.score) as rating
FROM movies m
JOIN ratings r ON m.movie_id = r.movie_id
GROUP BY m.movie_id, m.title
HAVING AVG(r.score) >= 8
ORDER BY rating DESC
LIMIT 5
\end{verbatim}

This semantic gap between natural language and formal queries represents the core challenge that this thesis addresses.

Furthermore, SQL's compositional nature allows complex queries to be built from simpler components through nesting and
combination of operations. NL2SQL systems must therefore not only fill in the individual query components but also how
to compose these to semantically accurate and syntactically correct query statements. This is particularly important when
dealing with complex natural language queries that may require multi-clause SQL statements involving subqueries, multiple
joins, and aggregation functions to answer.

\subsubsection{Normalization and Schema Design}

Database schemas typically follow normalization principles to eliminate redundancy and maintain data integrity. The normal forms
(1NF, 2NF, 3NF, BCNF) provide a set of design guidelines and rules for relational database design to reduce duplication,
inconsistencies and anomalies \citep{DatabaseSystemIntro}. Understanding normalization forms is crucial for NL2SQL systems to handle:

\begin{itemize}
    \item \textbf{Schema complexity} — Normalized schemas often distribute logically related information across multiple tables,
          requiring natural language interfaces to understand implicit relationships and generate appropriate joins.
    \item \textbf{Semantic mapping} — Humans typically think about data in denormalized, conceptual terms, while relational databases
          store data in its normalized form. NL2SQL systems must overcome this layout difference (e.g. there is no 1:1 mapping
          of concepts to tables).
    \item \textbf{Query complexity} — Retrieving simple information from relational databases may require multiple joins in normalized schemas,
          challenging NL2SQL systems to generate potentially complex SQL statements from simple user requests.
\end{itemize}

The tension between normalized database design and intuitive natural language use represents a key challenge that influences
the architecture and design decisions explored in subsequent sections of this thesis.