### 2025-10-01 thesis/thesis.tex

\documentclass{article}

\usepackage[a4paper]{geometry}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fancyhdr}
\usepackage{float}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{import}
\usepackage{titlesec}
\usepackage{xcolor}
\usepackage[natbibapa]{apacite}

\graphicspath{{./images/}}

\pagestyle{fancyplain}
\fancyhf{}
\lhead{\fancyplain{}{Mara Schulke} }
\rhead{\fancyplain{}{\today}}
\cfoot{\fancyplain{}{\thepage}}

\newcommand{\figuresource}[1]{
	\begin{center}Source: {#1}\end{center}
}

\titleformat{\section}{\normalfont\large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\normalfont\normalsize\bfseries}{\thesubsection}{1em}{}
\titleformat{\subsubsection}{\normalfont\small\bfseries}{\thesubsubsection}{1em}{}

\begin{document}

\import{title}{index}
\import{abstract}{index}
\import{toc}{index}

\newpage

% 1/3 Vorarbeit, Literature Review

\import{introduction}{index}

\newpage

\import{literature-review}{index}

\newpage

% 1/3 Theory, Decomp und System Design

\import{theoretical-foundations}{index}

\newpage

\import{design}{index}

% 1/3 Implementation und Evaluation
\import{implementation}{index}

\newpage

\import{evaluation}{index}

\newpage

\import{discussion}{index}

\newpage

\import{conclusion}{index}

\newpage

\import{appendices}{index}

\end{document}