### 2025-10-03 thesis/design/indexing.tex
Split and finish the sampling section

\subsection{Schema Indexing}

To choose the most relevant subset of samples from $\mathcal{S}$ for $k$-shot learning,
it is important that the SQL queries we supply are written for structurally similar SQL schemas
in order to minimize the structural difference between the supplied samples $\mathcal{S}_k$ and
the ground truth query $\omega_{ground}$ for a given natural language question.

For example $\omega_{ground}$ the question ``Give me all contacts for the user with the id 10''
might look very different depending on the database schema at hand, thus only taking into
account the similarity of the natural language question will yield inferior results.

\vspace{1em}

\begin{figure}[ht]
\hfill
\begin{minipage}[b]{0.45\linewidth}
\centering
\begin{verbatim}
CREATE TABLE users (
  id TEXT PRiMARY KEY
);

CREATE TABLE contacts (
  user_id TEXT NOT NULL,
  name TEXT NOT NULL,
  FOREIGN KEY (user_id) ON users(id),
  PRIMARY KEY (user_id, name)
);
\end{verbatim}
\caption{Normalized schema}
\end{minipage}
\hfill
\begin{minipage}[b]{0.35\linewidth}
\centering
\begin{verbatim}
CREATE TABLE users (
  id TEXT PRIMARY KEY,
  contacts TEXT[]
);
\end{verbatim}
\caption{Denormalized schema}
\end{minipage}
\hfill
\end{figure}

\vspace{1em}

% TODO: Finish this section on WWL graph kernels and distance calculation..

%is a set of id and distance tuples $\{(s_1, d_1), ..., (s_i,d_i)\}$
%where $s$ is the database schema name and $d$ is the distance to the
%\textit{current} schema. Thus a schema distance index has to be computed for each database schema
%\textsc{Natural} is running on.

%Thus.. $\mathcal{D}$