### 2025-10-06 thesis/implementation/index.tex
Write an outline for the impl section

\section{Implementation}

This chapter describes the practical implementation of the \textsc{Natural} system,
translating the formal design specifications into a working NL2SQL system.

\subsection{System Architecture}
% - Overall system design and component interaction
% - Pipeline implementation mapping to design functions (σ, φ, π, ρ, ν)
% - Data flow and processing stages
% - API design and interfaces

\subsection{Development Environment and Tools}
% - Programming languages and frameworks chosen
% - LLM infrastructure (OpenAI API, local models, etc.)
% - Database systems and connectors
% - Development tools and testing frameworks

\subsection{Embedding Space Construction}
% - Implementation of embedding function ι
% - Masking algorithm implementation for NL and SQL
% - Vector database setup and indexing
% - Training data preprocessing pipeline

\subsection{Schema Distance Indexing}
% - Graph representation implementation for database schemas
% - WWL kernel computation algorithms
% - Distance index construction and storage
% - Efficient similarity search implementation

\subsection{Core System Components}

\subsubsection{Example Selection ($\sigma$)}
% - Cosine similarity computation implementation
% - Multi-dimensional scoring (question, SQL, schema similarity)
% - Top-k retrieval algorithms
% - Performance optimization strategies

\subsubsection{Schema Subsetting ($$)}
% - SQL parsing for schema element extraction
% - Schema graph construction and pruning
% - Context window management
% - Schema serialization for LLM consumption

\subsubsection{Query Projection (π)}
% - LLM integration and prompt engineering
% - Few-shot prompt construction
% - Model selection and configuration
% - Response parsing and validation

\subsubsection{Self Refinement (ρ)}
% - Error detection and analysis systems
% - Iterative refinement loop implementation
% - Feedback integration mechanisms
% - Convergence criteria and stopping conditions

\subsubsection{Voting System (ν)}
% - Result comparison and frequency analysis
% - Consensus algorithms implementation
% - Tie-breaking strategies
% - Quality scoring mechanisms

\subsection{Whitelisting}\label{implementation:whitelisting}
% - Natural language keyword whitelist construction
% - Domain-specific vs. generic term handling
% - Empirical whitelist optimization
% - Cross-domain transferability considerations

\subsection{Performance Optimization}
% - Caching strategies for embeddings and distances
% - Parallel processing for candidate generation
% - Memory management for large schemas
% - Latency optimization techniques

\subsection{Error Handling and Robustness}
% - Input validation and sanitization
% - Graceful degradation strategies
% - Logging and monitoring implementation
% - Edge case handling

\subsection{Testing and Validation}
% - Unit testing for individual components
% - Integration testing for end-to-end pipeline
% - Performance benchmarking setup
% - Correctness validation frameworks

\newpage