### 2025-10-06 thesis/theoretical-foundations/graphs.tex
Add theory sections i need to write

\subsection{Graph Theory Fundamentals}

% TODO: Add graph theory basics section (needed for WWL kernel approach)
% - Basic graph definitions: vertices, edges, labeled graphs
% - Graph representations and adjacency structures
% - Graph similarity metrics and kernels
% - Weisfeiler-Lehman graph features
% - Wasserstein distance and optimal transport theory
% - Applications to database schema representation

% TODO: Connect to design section
% - How database schemas map to graph structures
% - WWL kernel computation for schema similarity
% - Distance indexing for efficient retrieval
% - Integration with example selection function Ã