### 2025-10-06 thesis/theoretical-foundations/index.tex

\section{Theoretical Foundations}

This section serves as an introduction to the theoretical concepts utilized in this thesis. It is meant to establish a
baseline of understanding for the design, implementation and evaluation phases. It briefly covers relational database theory
(incl. SQL), machine learning fundamentals and their subsequent application for large language models.

\subimport{}{relational-theory.tex}
\subimport{}{machine-learning.tex}
\subimport{}{nlp.tex}
\subimport{}{llms.tex}
\subimport{}{graphs.tex}
\subimport{}{gpu-computing.tex}

% TODO: Improve integration between theoretical sections
% - Add cross-references between related concepts
% - Show how machine learning enables LLMs
% - Connect graph theory to database schema representation
% - Link NLP techniques to practical system components

% TODO: Add forward connections to design section throughout theory
% - Reference specific design functions (σ, φ, π, ρ, ν) where relevant
% - Connect theoretical concepts to algorithmic implementations
% - Bridge gap between formal foundations and practical system design

%\subsection{Problem Decomposition}

%\subsection{Requirements}

% Conceptual Design

%\subsection{System Design}

%\subsubsection{Architecture Design}

%\subsection{Technical Implementation Strategies}