### 2025-10-06 thesis/theoretical-foundations/llms.tex

\subsection{Large Language Models}

% TODO: Complete LLM theoretical foundations section (essential for LLM-based approach)
% - Transformer architecture fundamentals
% - Attention mechanisms and self-attention
% - Pre-training vs fine-tuning paradigms
% - In-context learning and few-shot learning
% - Prompt engineering and instruction following
% - Model scaling laws and emergent capabilities
% - Limitations: hallucination, context windows, reasoning

% TODO: Connect to design section
% - How LLMs enable natural language understanding for SQL generation
% - Role in query projection function À
% - Self-refinement capabilities Á
% - Voting and consensus mechanisms ½