### 2025-10-06 thesis/theoretical-foundations/nlp.tex

\subsection{Natural Language Processing}

% TODO: Complete NLP theoretical foundations section
% - Text preprocessing and tokenization
% - Word embeddings and semantic representations
% - Language modeling fundamentals
% - Syntactic and semantic parsing
% - Named entity recognition and linking
% - Text similarity metrics (cosine, semantic distance)
% - Masked language modeling approaches

% TODO: Connect to design section
% - Role in natural language query understanding
% - Embedding space construction for example selection
% - Masking techniques for domain generalization
% - Semantic similarity computation in function Ã