### 2025-11-08 thesis/appendices/prompts.tex

\subsection{Prompts}

\subsection{Inference Prompt}\label{appendix:inference-prompt}

\begin{verbatim}
EXAMPLE OF THE PROMPT
\end{verbatim}

\subsection{Refinement Prompt}\label{appendix:refinement-prompt}

\begin{verbatim}
EXAMPLE OF THE PROMPT
\end{verbatim}