### 2025-11-16 thesis/theoretical-foundations/index.tex
Rewrite the theory section intro

\section{Theoretical Foundations}

This section introduces the theoretical concepts needed for understanding
\textsc{Natural}'s design and implementation sections. It covers relational
database theory, machine learning fundamentals, large language models, natural
language processing, embeddings and graph theory.

These foundations directly support the system's core functions: example
selection ($\sigma$), schema subsetting ($\phi$), query projection ($\pi$),
self-refinement ($\rho$), and voting ($\nu$).

\subimport{}{relational-theory.tex}
\subimport{}{machine-learning.tex}
\subimport{}{llms.tex}
\subimport{}{nlp.tex}
\subimport{}{graphs.tex}