### 2025-11-16 thesis/theoretical-foundations/machine-learning.tex
Write ml section

\subsection{Machine Learning Fundamentals}

Machine learning represents the algorithmic foundation for state-of-the-art
NL2SQL systems, enabling systems to learn how natural language queries
translate into SQL. Understanding these foundations is essential for
implementing and optimizing large language model-based NL2SQL approaches.

\subsubsection{Neural Network Architecture}

Neural networks are the computational concept behind contemporary NL2SQL
approaches, representing complex capabilities through compositions of simpler
operations. A neural network is comprised of interconnected computational units
(called neurons) organized in layers, where each connection has an associated
(and learnable) weight and bias parameters.

The fundamental computation paradigm of neural network outputs is called
forward propagation, which is applying the respective weights ($W$) and biases
($b$) to the input parameter ($a$) and transforming it using a (non-linear)
activation function $f$:

\begin{equation}
a^{l+1} = f(Wa^l + b)
\end{equation}

where $W \in \mathbb{R}^{m \times n}$ is the weight matrix, $a \in
\mathbb{R}^{n}$ is the input vector, $b \in \mathbb{R}^{m}$ is the bias vector,
and $f$ is an activiation function (e.g. ReLU, sigmoid, or tanh). Using this
function repeatedly, propagates information through the neural network.

The most important architectures for simple natural language
processing networks include:

\begin{itemize}
    \item \textbf{Feedforward Networks} — Which can process fixed-size inputs
        through successive linear transformations and activations, this
        architecture is particularly suitable for classification and regression
        problems within NL2SQL systems.
    \item \textbf{Recurrent Networks (RNNs/LSTMs)} — Which can handle
        sequential data of variable-length by maintaining hidden states that
        capture dependencies, enabling processing of natural language sequences
        of arbitrary length.
    \item \textbf{Embedding Layers} — Which map discrete tokens (e.g. words,
        characters, or subwords) to a dense vector representation, thus
        providing the foundation for subsequent neural language processing.
\end{itemize}

The ability of neural networks to learn hierarchical representations through
multiple layers makes them particularly well-suited for the complex translation
required in NL2SQL systems.

\subsubsection{Learning Approaches}

Broadly speaking, there are three fundamental approaches to training a machine
learning model, also known as the process of ``learning'':

\begin{enumerate}
    \item \textbf{Supervised Learning} – The process of training a model
        using a labeled dataset of input and output pairs. This learning
        approach is quite commonly used in models used for NL2SQL which were
        trained on question and answer pairs (natural language question and
        corresponding SQL query). This learning approach aims to minimise the
        loss between predicitions and ground truth.
    \item \textbf{Unsupervised Learning} – Includes learning patterns that work
        with unlabeled datasets and use approaches like clustering, language
        modeling and heuristics to infer structure and yield predictions.
        Unsupervised learning is most appropriate when there is no clear,
        singular answer and the domain provides large datasets that are not
        feasible to label manually (eg, dataset gathered through webscraping
        the internet).
    \item \textbf{Reinforcement Learning} – The process of learning through
        reward signals, commonly known as the reward function. Applied to
        NL2SQL systems this could transfer to training a model's SQL generation
        capabilities based on execution feedback. Historically this learning
        pattern is not widely used in NL2SQL approaches.
\end{enumerate}

\subsubsection{Attention Mechanisms}

Attention mechanisms allow machine learning models to selectively focus on
relevant parts of their input by learning to weight different sections of the
input sequence based on its relevance to the current computation. While RNNs
processed sequences sequentially, attention enables direct modeling of
relationships between any positions in a sequence, which enabled the
transformer architecture to become prevalent for natural language processing
\citep{AttentionIsAllYouNeed}.

Attention can be computed by retrieving the weighted sum of values $V$
based on the compatibility between queries $Q$ and keys $K$:

\begin{equation}
\text{Attention}(Q, K, V) = \text{softmax}\left(\frac{QK^T}{\sqrt{d_k}}\right)V
\end{equation}

where $d_k$ is the dimensionality of the key vectors, used for scaling to
prevent softmax saturation which would incurr small gradients. The softmax
operation converts compatibility scores into a probability distribution,
determining how much attention to pay to each value.

This concept proved to be a foundational advancement to transformer
architectures seen in natural language processing and subsequently recent
NL2SQL research. Transformers extended the concept of attention into
self-attention and multi-head attention patterns that underpin todays large
language models.

\subsubsection{Transformer Architecture}

The transformer architecture, introduced by \citeauthor{AttentionIsAllYouNeed}
in \citeyear{AttentionIsAllYouNeed}, introduced the model architecture found in 
modern large language models \citep{AttentionIsAllYouNeed}. Unlike earlier
sequence to sequence models, transformers process are capable of processing
entire sequences of inputs in parallel by using a self-attention mechanism.
This enabled more efficient training on massive datasets and superior
representation of of complex dependencies in input sequences.

\paragraph{Self-Attention Mechanism}

Self-attention allows each position in a sequence to be weighted to all other
positions in the sequence, thus resulting in a representations that can capture
contextual relationships between input tokens. For an input sequence of tokens
$x_1, \ldots, x_n$, the self-attention mechanism computes:

\begin{equation}
\text{SelfAttention}(X) = \text{softmax}\left(\frac{QK^T}{\sqrt{d_k}}\right)V
\end{equation}

where $Q = XW_Q$, $K = XW_K$, and $V = XW_V$ are learned linear projections of
$X$ into the query, key and value representations. This mechanism enables
transformers to weigh the importance of different tokens while encoding each
position. This maintains semantic and syntactic relationships which are crucial
for understanding natural language queries in NL2SQL systems.

\paragraph{Multi-Head Attention}

Multi-head attention extended the self-attention concept by computing multiple
attention operations in parallel, allowing the model to put focus to different
aspects of the input at the same time:

\begin{equation}
\text{MultiHead}(Q, K, V) = \text{Concat}(\text{head}_1, \ldots, \text{head}_h)W^O
\end{equation}

where each $head$ computes:

\begin{equation}
\text{head}_i = \text{Attention}(QW_i^Q, KW_i^K, VW_i^V)
\end{equation}

Each attention head can specialize in capturing a different type of semantic
relationship. For example one head might focus on syntactic structure (e.g., linking SQL
keywords to their arguments), while others might capture semantic relationships
between works (e.g., List all x where y). This multi-attention mechanism is
essential for the complex tasks which require capturing multiple different
semantic relationships between words.