### 2025-11-16 thesis/theoretical-foundations/llms.tex
Extend the llm theory

\subsection{Large Language Models}

Large language models (LLMs) are a recent advancement built on top of the
transformer architecture for natural language processing, demonstrating state
of the art capabilities in understanding, reasoning and text generation. These
models form the core of most modern NL2SQL systems like \textsc{XiYan-SQL},
\textsc{DAIL} and others. They have a wide language understanding as they have
been trained on large text corpora, are exposed to different domains of traning
data and have an inherent understanding of the SQL syntax as they have been
exposed to it during pre-training. This shifted the approach while developing
NL2SQL systems to focus on in-context learning methods instead of training
dedicated models from scratch.

\subsubsection{Feedforward and Residual Connections}

Every layer in a transformer combines a multi-head attention with position-wise
feedforward networks:

\begin{equation}
\text{FFN}(x) = \text{GELU}(xW_1 + b_1)W_2 + b_2
\end{equation}

Residual connections and layer normalization stabilize training of complex networks:

\begin{equation}
    \text{Output} = x + \text{Sublayer}(\text{LayerNorm}(x))
\end{equation}

where $\text{Sublayer}$ represents either multi-head attention or the
feedforward network. These components enable transformers to scale to billions
of parameters while maintaining training stability.

\subsubsection{Positional Encoding}

Since transformers are able to process sequences in parallel, positional
encodings inject information about token positions. Original transformers use
sinusoidal functions, while modern LLMs often employ learned positional
embeddings or relative position encodings. This positional information is
inherently important for NL2SQL systems, where the order of generated SQL
clauses (e.g., \texttt{SELECT} must come before before \texttt{WHERE}) carries
significant semantic meaning.

\subsubsection{Pre-training and Fine-tuning}

Large language models are trained in a two-fold process that separates the
training of general language understanding from from task-specific
capabilities.

\paragraph{Pre-training}

During pre-training LLMs are exposed to vasts amounts of unlabeled text
gathered from the internet, books etc. Using self-supervised objectives, no
manual labeling of the datasets is required. The most common pre-training
objective is causal language modeling, where a model learns to predict the next
token given a previous context:

\begin{equation}
\mathcal{L}_{\text{LM}} = -\sum_{t=1}^{T} \log P(x_t \mid x_{<t}; \theta)
\end{equation}

This objective engrains general language patterns, common knowledge,
reasoning capabilities, and even programming knowledge into models 
when code is included in the training datasets. Models like GPT, LLaMA, and Qwen
are pre-trained using this objective, aiming to predict coherent
continuations of text sequences.

\paragraph{Fine-tuning}

Fine-tuning is the process of adapting a pre-trained model, for example for
solving NL2SQL tasks, by continuing model training on a labeled dataset
(eg, natural language question to expected sql query). During fine-tuning, the
model can learn to map natural language questions to expected SQL queries when
being presented with a question and a database schema.

Fine-tuning usualy require significantly smaller datasets (thousands to
hundreds of thousands of examples) compared to pre-training (billions to
trillions of tokens), largely relying on the natural language understanding
capabilities acquired during the pre-training phase.

Models like \textsc{OmniSQL} demonstrate that smaller fine-tuned models 
can outperform significantly larger LLMs for NL2SQL tasks using a fraction of
the parameters \citep{OmniSQL}.

\subsubsection{In-Context Learning}

In-context learning (ICL) is the process of teaching a model how to perform a
certain task by providing examples in its prompt, without updating models
parameters. Thus ICL and Fine-tuning differ by their timing: runtime vs learning.

This capability emerged with sufficiently large models which have large enough
context windows to fit in examples of solutions for similar problems. This
method was shown to be very effective by \citeauthor{DAIL} in \citeyear{DAIL},
improving the performance of general purpose LLMs as well as fine-tuned LLMs
when being presented with relevant examples \citep{DAIL}. The combination of
fine-tuned models like \textsc{OmniSQL} with ICL is therefore a promising
research for NL2SQL systems.

\subsubsection{Prompt Engineering}

Prompt engineering is the process of structuring the input to LLMs in a way
where they adhere to desired behaviours. For NL2SQL systems previous research
has found diverging effectiveness between different example, schema and
question presentation mechanisms \citep{DAIL, OmniSQL}.

\subsubsection{Chain-of-Thought Reasoning}

Prompting or training models to output their chain-of-thought encourages models 
to generate reasoning traces before producing their final answer to a question
or input. For NL2SQL, this usually involves generating explanations of query
logic, analysing the input database schema, and planning the query structure
before generating a query.

Fine-tuned models like \textsc{OmniSQL} are fine-tuned to produce
chain-of-thought output during generation in markdown which enables debugging
model behaviour and understanding the query generation approach.

\subsubsection{Model Limitations}

Understanding the limitations of large language models is essential for
designing robust NL2SQL systems.

\paragraph{Hallucination}

As LLMs are just predicting a sequence of output tokens, they have no inherent
understanding of the actual problem domain and solution space they are
presented with. Thus in NL2SQL systems, LLMs might generate invalid SQL queries
which might reference non existent tables, columns or relationships, output
invalid syntax or generate promising but semantically invalid queries.

These flaws are inherent to a text based representation and are non
recoverable, although through candidate validation and self-refinement
approaches most obviously invalid queries can be recovered. Detecting semantic
invalidity is a hard research problem on its own.

\paragraph{Context Window Limitations}

As the underlying transformer architecture is still bound to finite context
windows, extensive database schemas or long lists of examples can exceed the
window, thus causing the query generation to fail.