### 2025-11-23 thesis/theoretical-foundations/nlp.tex
Write about nlp

\subsection{Natural Language Processing}

Natural language processing (NLP) provides the foundations for working with
natural language inputs in computer science and mathematics. It allows for text
understanding and text representation, aswell as text processing and semantic
feature extraction, capturing meaning and structure. For NL2SQL systems NLP
enable semantic understanding of natural language queries and semantic search
using linguistic similarity. Using NLP and text embeddings reference datasets
can be used for semantic search effectively and accurately.

\subsubsection{Text Preprocessing and Tokenization}

Before neural networks can process text, raw text must be converted into a
structured representations that models can process. This transformation is called
tokenization, which refers to the process of segmenting text into discrete
units (tokens) for subsequent processing.

\paragraph{Tokenization Approaches}

Tokens are short character chains that allow representing text efficiently as a
series of tokens by refering to the tokenizers vocabulary. Therefore NLP
systems mostly employ subword tokenization. Modern tokenization approaches
balance vocabulary size with representational efficiency. Common approaches
include \textbf{Byte-Pair Encoding (BPE)}, \textbf{WordPiece} and
\textbf{SentencePiece}.

These subword tokenization approaches are particularly valuable for NL2SQL
because database schemas often contain domain-specific terminology, compound
words, and technical terminology that may not appear in common vocabularies.
Subword approaches handle unseen words gracefully by decomposing them into
meaningful tokens.

\paragraph{Tokenization in NL2SQL Contexts}

For NL2SQL systems, tokenization must be capable of both encoding natural
language queries and SQL queries. SQL presents a challenge in NLP as it relies
on a structured syntax and usage of special characters (dots, underscores,
SQL operators, etc.) which are less commonly found in natural language.

\subsubsection{Word Embeddings and Semantic Representations}

Word embeddings are the vector space representation of a tokens where semantic
meaning is captured through geometric properties. As word embeddings are
ordinary vectors in a high-dimensional vector space, common vector operations
can still be performed.

\paragraph{Embedding Functions}

An embedding function maps tokens from a discrete vocabulary $V$ to dense
vectors in $\mathbb{R}^d$:

\begin{equation}
e_{w}: V \rightarrow \mathbb{R}^d
\end{equation}

where $d$ is the embedding dimensionality (usually between 256-8192 for modern
models). The key property of the vector space $\mathbb{R}^d$ is that
semantically similar words occupy nearby regions in this vector space, hence
vector operations and proximity in the vector space allows for computation of
semantic properties such as semantic distance.

Early embedding methods like Word2Vec and GloVe learned static representations
where each word has a single vector regardless of its surrounding context.
Modern contextual embeddings, generate different vectors for the same word
based on its context, thus capturing meaning more accurately.

\paragraph{Sentence Embeddings}

While word embeddings are useful for determining the relationships between
words, they are limited in length and complexity. NL2SQL systems require
representations of entire questions and SQL queries in order to discover
answers to similar questions asked in the past. Sentence embeddings aggregate
the token-level meaning into fixed-size vectors in the same vector space
$\mathbb{R}^d$, thus representing complete sentences or texts in one vector.

\begin{equation}
    e_{s}: (v_0, .., v_n) \rightarrow \mathbb{R}^d
\end{equation}

Different aggregation / pooling strategies emerged, like mean pooling, max
pooling, using a CLS token or learned aggregation. \textsc{Natural} uses
an embedding model (Qwen3-Embedding-8B) trained specifically for generating
high-quality sentence representations for subsequent searching.

\subsubsection{Semantic Similarity Metrics}

Quantifying the similarity between two natural language sentences is essential
for example selection in NL2SQL. This requires that similarity metrics in the
vector space align with natural understanding of semantic relation.

\paragraph{Cosine Similarity}

Cosine similarity is a measurement using the angle between two vectors,
to capture their directional similarity while normalizing for magnitude:

\begin{equation}
s_{\cos} = \frac{u \cdot v}{\|u\| \|v\|} = \frac{\displaystyle\sum_{i=1}^n u_i v_i}{\sqrt{\displaystyle\sum_{i=1}^n u_i^2} \sqrt{\displaystyle\sum_{i=1}^n v_i^2}}
\end{equation}

Thus cosine similarity ranges from -1 which indicates opposite directions to 1
representing same direction, with 0 indicating orthogonality of the input
vectors. This metric is widely used in NLP as it is independent of vector
magnitude, efficient to compute and well-suited for high-dimensional vector
spaces.

\paragraph{Cosine Distance}

Cosine distance converts similarity into a distance metric:

\begin{equation}
d_{\cos}(u, v) = 1 - \text{s}_{\cos}(u, v)
\end{equation}

Thus cosine distance ranges from 0 for identical direction to 2 for opposite
direction input vectors. This is satisfying properties desirable for distance
metrics while preserving the angular relationship of the input vectors $u$ and
$v$.