### 2025-11-23 thesis/discussion/index.tex
Benchmarking section outline

% Discussion
\section{Discussion}

\subsection{Discussion of Findings}

% TODO: Introduction paragraph synthesizing evaluation results

\subsubsection{Research Questions Revisited}

\paragraph{RQ1: Feasibility for Real-World Application}
% TODO: Answer RQ1 directly with evidence
% - Is semantic accuracy high enough? (cite Spider/Bird EA from 7.2)
% - Can it run on reasonable hardware? (cite performance analysis from 7.4)
% - Overall verdict: Yes/No/Qualified, with supporting evidence

\paragraph{RQ1.1: Semantic Accuracy Threshold}
% TODO: Address sub-question on accuracy benefits
% - What EA threshold provides "noticeable benefit"?
% - How does achieved EA compare to this threshold?
% - User tolerance for errors in different contexts

\paragraph{RQ1.2: Hardware Accessibility}
% TODO: Address sub-question on hardware requirements
% - RTX 3090 (24GB VRAM): Is this "mass-available"?
% - Comparison to high-end research GPUs (A100, H100)
% - Could system run on lower-tier GPUs with optimizations?

\paragraph{RQ2: Ambiguity Resolution Approaches}
% TODO: Answer RQ2 with evidence from ablations
% - Which components most effectively resolve ambiguity? (cite ablation study 7.3)
% - Ranking: σ, φ, ρ, ν by impact on ambiguous queries
% - Context-dependent effectiveness (different approaches for different ambiguity types)

\paragraph{RQ3: Strategies for Increasing Semantic Accuracy}
% TODO: Answer RQ3 with evidence from ablations
% - Which strategies provide largest accuracy gains? (cite 7.3.6 cumulative analysis)
% - Are gains cumulative or synergistic?
% - Optimal combination of techniques
% - Comparison to literature: do findings align with prior work?

\subsubsection{Real-World Viability Assessment}

\paragraph{Practical Deployment Considerations}
% TODO: Discuss deployment feasibility
% - Integration with PostgreSQL architecture
% - Deployment as extension vs external service
% - Security considerations (SQL injection risks from LLM output)
% - Privacy considerations (schema/data exposure to model)

\paragraph{User Experience Implications}
% TODO: Discuss UX trade-offs
% - Latency acceptable for interactive use?
% - How to communicate uncertainty to users?
% - Handling ambiguous queries: clarification dialogs?
% - Trust and transparency: showing generated SQL to users

\paragraph{Comparison to Traditional SQL Interfaces}
% TODO: Position system relative to traditional approaches
% - When is NL interface better than SQL?
% - When should users still write SQL manually?
% - Complementary use cases (NL for exploration, SQL for production)
% - Learning curve: does NL interface help SQL learning or hinder it?

\paragraph{Accessibility and Democratization Benefits}
% TODO: Discuss broader impact
% - Does system truly make databases more accessible?
% - Who benefits most? (Non-technical users, analysts, data scientists)
% - Limitations for complex analytical queries
% - Role in data democratization initiatives

\subsubsection{Limitations and Threats to Validity}

\paragraph{Benchmark Limitations}
% TODO: Discuss external validity threats
% - Spider/Bird may not represent all real-world queries
% - Academic benchmarks vs production query distributions
% - Domain coverage gaps
% - Temporal limitations (benchmarks are static snapshots)

\paragraph{Model-Specific Results}
% TODO: Discuss generalizability to other models
% - OmniSQL 7B is one specific model
% - Would results hold for other LLMs? (GPT-4, Claude, Llama, Qwen)
% - Fine-tuning specificity: OmniSQL trained on Spider/Bird
% - Reproducibility with future model generations

\paragraph{Hardware-Specific Results}
% TODO: Discuss generalizability to other hardware
% - RTX 3090 specific (not A100, not Apple Silicon)
% - CUDA-specific optimizations
% - Would results translate to other GPU architectures?
% - Cloud deployment characteristics may differ

\paragraph{Evaluation Metric Limitations}
% TODO: Discuss measurement validity threats
% - EA doesn't capture semantic equivalence perfectly
% - EM is too strict (many equivalent SQL formulations)
% - Missing metrics: query readability, maintainability, efficiency
% - User satisfaction not measured

\paragraph{Comparison Validity}
% TODO: Discuss comparison fairness
% - Different SOTA systems use different models, hardware
% - Not apples-to-apples comparisons
% - Cherry-picking risks in literature
% - Reproducibility challenges

\paragraph{Internal Validity Considerations}
% TODO: Discuss experimental design limitations
% - Ablation study design choices
% - Hyperparameter tuning: were optimal settings found?
% - Statistical power: sample sizes sufficient?
% - Confounding factors

\subsubsection{Implications for Future Work}

\paragraph{System Improvements}
% TODO: Recommend concrete system enhancements
% - Components that deserve deeper investigation
% - Alternative approaches to explore (e.g., different schema subsetting strategies)
% - Performance optimizations (caching, batching, quantization)
% - Integration of newer LLMs

\paragraph{Evaluation Enhancements}
% TODO: Recommend additional experiments
% - User studies with real users (not just benchmark metrics)
% - Longitudinal deployment studies
% - Domain-specific fine-tuning experiments
% - Alternative benchmark datasets
% - Comparative studies with commercial tools (e.g., GitHub Copilot for databases)

\paragraph{Theoretical Contributions}
% TODO: Discuss research directions
% - Formalization of ambiguity resolution in NL2SQL
% - Better understanding of component synergies
% - Transfer learning from NL2SQL to other semantic parsing tasks
% - Formal verification of generated SQL queries

\paragraph{Practical Deployment Research}
% TODO: Recommend real-world validation
% - Pilot deployments in production environments
% - Integration with BI tools (Tableau, Power BI)
% - Education contexts (teaching SQL through NL)
% - Accessibility studies (impact on non-technical users)

\newpage