### 2025-11-23 thesis/evaluation/index.tex

% Reference: https://github.com/petavue/NL2SQL-Benchmark/blob/main/results/Report/NL2SQL%20Benchmark%20Report.pdf

\section{Evaluation}

The evaluation phase of \textsc{Natural} builds upon the benchmarking
infrastructure outlined in section~\ref{impl:benchmark}. This section is
evaluating benchmarking data gathered while running \textsc{Natural} in
different configurations on two prevalent benchmarking datasets:
\textsc{Spider} and \textsc{Bird}. A representative baseline is introduced,
aiming to mirror the base-model performance. Subsequently an ablation study is
performed to understand the performance impact of individual pipeline components.

\subsection{Experimental Methodology}

% TODO: Introduction paragraph explaining the systematic evaluation approach

\subsubsection{Test Environment}

The tests are performed on a consumer-grade linux system, as research- or
enterprise-grade hardware was not available. Thus only a subset of datasets and
configurations have been measured, as a single benchmark executions on prevalent
datasets takes 12-36 hours.

\paragraph{Hardware Configuration}

The system used for benchmarking has a NVIDIA RTX 3090 GPU with 24GB VRAM
available, an AMD 9990X3D CPU with 12-cores up to 5.5GHz, 64GB of RAM at
6400mt/s and 1TB of SSD storage at a read speed of 7450 MB/s and a write speed
of 6900 MB/s.

\paragraph{Software Stack}

The software stack running on the system used for benchmarking is NixOS (25.05)
with a recent linux kernel version (6.17.7). CUDA 13.0 is being used for
inference on GPUs and a rust compiler version is 1.91 for compiling
\textsc{Natural}. The models used during benchmarking are \textsc{OmniSQL 7B
Q8\_0} and \textsc{Qwen3 Embedding 8B Q4\_K\_M} for generation and embedding
respectively.

\paragraph{System Configuration}

% TODO: Describe configuration parameters
% - Model loading settings
% - Batch sizes
% - Timeout configurations
% - Any other relevant parameters

\subsubsection{Benchmark Datasets}

The benchmarking datasets used are the \textsc{Spider} and \textsc{Bird}
testdatasets that are prevalently used for evaluation of NL2SQL systems.
Comparable research has frequently published performance metrics
(\textsc{Execution Accuracy} and \textsc{Exact Match}) on these benchmarks
which makes the performance of \textsc{Natural} comparable to other works.

\paragraph{Spider Test Set}

The \textsc{Spider}

% TODO: Describe Spider benchmark
% - Dataset size: 10,181 questions, 5,693 SQL queries
% - 200 databases across 138 domains
% - Difficulty distribution (Easy/Medium/Hard/Extra Hard)
% - Why Spider was chosen
% - Train/dev/test split strategy

\paragraph{Bird Benchmark}
% TODO: Describe Bird benchmark
% - Dataset size: 12,751 pairs across 95 databases (33.4 GB)
% - 37 domains
% - Real-world characteristics (dirty data, external knowledge requirements)
% - Why Bird was chosen as complementary to Spider

\paragraph{Dataset Preprocessing}
% TODO: Describe any preprocessing steps
% - Database preparation
% - Schema extraction
% - Gold standard SQL formatting

\subsubsection{Evaluation Metrics}

\paragraph{Execution Accuracy (EA)}
% TODO: Define execution accuracy
% - Definition: whether query results match ground truth
% - Calculation method
% - Why EA is the primary metric
% - Reference to natural-benchmark implementation

\paragraph{Exact Match (EM)}
% TODO: Define exact match
% - Definition: whether SQL query exactly matches ground truth
% - Calculation method
% - Limitations and interpretation

\paragraph{Performance Metrics}
% TODO: Define performance metrics
% - End-to-end latency (time from NL query to SQL result)
% - Per-component latency (σ, φ, π, ρ, ν timings)
% - Throughput (queries per second)
% - Resource utilization (GPU memory, CPU usage)

\paragraph{Statistical Significance}
% TODO: Describe statistical testing approach
% - Test selection (e.g., paired t-test, McNemar's test)
% - Significance threshold (p < 0.05)
% - Confidence intervals

\subsubsection{Baseline Systems and Comparisons}

\paragraph{Internal Baselines}
% TODO: Describe ablated baselines
% - Zero-shot baseline (π only, no σ, φ, ρ, ν)
% - Partial system variants for ablation studies

\paragraph{State-of-the-Art Comparisons}
% TODO: List SOTA systems for comparison
% - GPT-4 results from literature
% - DIN-SQL, DAIL-SQL, C3 results
% - Other relevant open-source approaches
% - Note: Direct comparison constraints (different models, hardware)

\subsection{Benchmark Performance Analysis}

% TODO: Introduction paragraph presenting overall findings

\subsubsection{Spider Results}

\paragraph{Overall Performance}
% TODO: Present overall EA and EM scores
% - Table: Overall Spider dev set results
% - Comparison to baseline
% - Comparison to SOTA from literature

\paragraph{Performance by Difficulty Level}
% TODO: Break down results by Spider difficulty classification
% - Table: EA/EM by Easy/Medium/Hard/Extra Hard
% - Analysis of performance degradation with difficulty
% - Discussion of which difficulty levels benefit most from system components

\paragraph{Performance by Database}
% TODO: Present per-database analysis
% - Identify databases where system excels
% - Identify databases where system struggles
% - Analyze characteristics that correlate with performance

\paragraph{Performance by SQL Complexity}
% TODO: Analyze results by SQL pattern complexity
% - Simple SELECT queries
% - JOIN operations
% - Nested subqueries
% - Aggregations and GROUP BY
% - Set operations (UNION, INTERSECT, EXCEPT)
% - Complex predicates

\paragraph{Learning Curve Analysis}
% TODO: Analyze impact of training example quantity
% - How does performance change with k-shot examples?
% - Diminishing returns analysis
% - Optimal k value determination

\subsubsection{Bird Results}

\paragraph{Overall Performance}
% TODO: Present overall Bird results
% - Table: Overall Bird dev set results
% - Comparison to baseline
% - Comparison to SOTA from literature (GPT-4: 54.89%, human: 92.96%)

\paragraph{Domain-Specific Performance}
% TODO: Break down by Bird's 37 domains
% - Best-performing domains
% - Worst-performing domains
% - Domain characteristics that impact performance

\paragraph{Dirty Data Handling}
% TODO: Analyze performance on Bird's dirty data cases
% - How does system handle real-world data quality issues?
% - Error patterns related to data quality
% - Comparison to clean data performance

\paragraph{External Knowledge Requirements}
% TODO: Analyze cases requiring external knowledge
% - Identify queries needing domain knowledge
% - System performance on these cases
% - Limitations of pure schema-based approaches

\subsubsection{Comparison with State-of-the-Art}

\paragraph{Quantitative Comparison}
% TODO: Create comprehensive comparison table
% - Table: System vs GPT-4, DIN-SQL, DAIL-SQL, C3, etc.
% - Both Spider and Bird results
% - Note differences in model architecture and hardware

\paragraph{Open-Source vs Proprietary Models}
% TODO: Discuss open-source (OmniSQL 7B) vs proprietary (GPT-4) trade-offs
% - Performance gap analysis
% - Cost and accessibility advantages
% - Privacy and control benefits
% - Deployment feasibility differences

\paragraph{Qualitative Comparison}
% TODO: Beyond metrics, discuss systemic differences
% - System architecture differences
% - Component strategies (how others handle schema subsetting, refinement, etc.)
% - Integration and deployment characteristics

\subsection{Performance Characteristics}

% TODO: Introduction paragraph on performance evaluation methodology

\subsubsection{Latency Analysis}

\paragraph{Per-Component Timing Breakdown}
% TODO: Present timing for each pipeline component
% - Table: Average latency for σ, φ, π, ρ, ν
% - Distribution statistics (median, p50, p95, p99)
% - Identify bottlenecks

\paragraph{End-to-End Query Response Time}
% TODO: Present total pipeline latency
% - Distribution of total response times
% - Comparison to acceptable thresholds (what's "interactive"?)
% - Impact of query complexity on latency

\paragraph{Latency vs Accuracy Trade-offs}
% TODO: Analyze speed-accuracy Pareto frontier
% - Can we reduce latency by limiting refinement iterations?
% - Can we reduce latency by using fewer candidates for voting?
% - Configurable performance profiles (fast mode vs accurate mode)

\paragraph{Bottleneck Identification and Optimization Opportunities}
% TODO: Identify where time is spent
% - Model inference time (largest component?)
% - Embedding search time
% - Graph kernel computation time
% - Database execution time (for refinement)
% - Recommendations for optimization

\subsubsection{Resource Utilization}

\paragraph{GPU Memory Usage}
% TODO: Analyze VRAM consumption
% - Peak memory usage during inference
% - Memory required for model loading
% - Impact of schema size on memory (large schemas → CUDA OOM)
% - Memory optimization strategies employed

\paragraph{CPU and System Memory}
% TODO: Analyze CPU usage patterns
% - CPU utilization during pipeline execution
% - RAM consumption for embeddings, candidate storage
% - I/O patterns (database access, vector index reads)

\paragraph{Disk I/O and Vector Index Performance}
% TODO: Analyze storage subsystem impact
% - SQLite-vec index size
% - Read performance for similarity search
% - Impact of index size on query latency

\paragraph{Power Consumption and Thermal Characteristics}
% TODO: If measured, discuss energy efficiency
% - GPU power draw during inference
% - Comparison to CPU-only approaches
% - Sustainability considerations

\subsection{Qualitative Analysis}

% TODO: Introduction paragraph explaining value of qualitative evaluation beyond metrics

\subsubsection{Case Studies}

\paragraph{Success Cases}
% TODO: Present 3-5 exemplar success cases
% - Full trace: NL query → schema → examples → generated SQL → execution results
% - Analysis of what made these queries succeed
% - Demonstrate system capabilities at their best
% - Show component contributions in action

\paragraph{Failure Cases}
% TODO: Present 3-5 exemplar failure cases
% - Full trace showing where system failed
% - Root cause analysis: schema linking error? Semantic error? Syntactic error?
% - What component(s) failed?
% - Could failure be prevented with different design choices?

\paragraph{Edge Cases and Boundary Conditions}
% TODO: Present interesting edge cases
% - Very ambiguous queries
% - Queries requiring domain knowledge
% - Queries with multiple valid interpretations
% - Queries at the limit of SQL expressiveness
% - How system behaves at boundaries

\newpage