### 2025-12-14 thesis/title/index.tex
Title

\begin{titlepage}
    \begin{center}
        \begin{Large}
            Brandenburg University of Applied Sciences \\[1em]
        \end{Large}
        IT Security \\
        Computerscience \\
        Prof. Dr. Oleg Lobachev \\
        MSc. Florian Eich
    \end{center}

    \vfill

    \begin{center}
        \Large{Applied Graph Kernels for Schema-Aware In-Context Learning in NL2SQL}\\[0.5em]
        \large{Bachelor Thesis}\\[1em]

        \begin{normalsize}
            Winter Semester 2025\\[0.25em]
            \today
        \end{normalsize}
    \end{center}

    \vfill

    \begin{center}
        Mara Schulke – Matr-Nr. 20215853
    \end{center}
\end{titlepage}